Supplementary Materials

### 2023-07-01 manuscript/manuscript.tex
supp, methods
more methods

(c.f. \cite{Robertson1961-ho,Santiago1995-hx}; see Supplementary Section
\ref{suppsec:theory} for a proof). In general, however, statistics such as
heterozygosity depend on the cumulative autocorrelation factor up to some time
$t$, $Q_t^2$ \parencite[p. 2111]{Santiago1998-bs}. This reflects the fact that
effective population size experienced by a mutation newly associated with a
fitness background is larger than the asymptotic effective population size,
since not much autocorrelation has accumulated in the stochastic perturbations
caused by drift yet. From a backwards-time perspective, this reflects the fact
that pairwise diversity (which is approximately heterozygosity when $2N\mu \ll
1$) depends on the pairwise coalescence rate per generation, which is not
constant under draft. The benefit of using Robertson's forward-time model of
draft is that the inflation factor is invariant with respect to the particular
fitness background the neutral allele becomes stochastically associated with.
By contrast, the difficulty with modeling draft backwards in time is that the
coalescence rates experienced by a lineage are not invariant to which lineage
was sampled due to its particular associated fitness background.

@@

where the factor of one-half comes from ignoring weak associations from
unlinked regions and chromosomes (see Supplementary Section
\ref{suppsec:heritable-fitness}). In our genome-wide model, we consider the
summation in Equation \eqref{eq:polygenic_Ne} over non-overlapping segments $i
\in \{1, 2, \ldots, S\}$ each undergoing selection such that segment $i$
contributes additive fitness variance $V_{A,i}$ to the total additive genetic
fitness variance. Under equilibrium levels of fitness variation, the
autocorrelation function for a neutral allele associated with segment $i$ is
$C(t) = [(1-r_i)(1-\kappa_i)]^t$, where $r_i$ is the recombination fraction to
the segment and $\kappa_i$ is the rate that the associated fitness variance
decays due to selective dynamics. Then, the cumulative autocorrelation is,

@@

Similar to previous work, our method characterizes the negative selection
process through the signal left on pairwise diversity at linked sites. Although
pairwise diversity is a less-informative statistic than the full site frequency
spectrum, it is much easier to derive models of how selection alters expected
pairwise coalescence times at linked sites than to describe the entire
genealogy. Spatial patterns of pairwise diversity are highly variable across
eukaryotic genomes, as linked selection leads diversity to become correlated
with spatial heterogeneity in recombination rates and the density of conserved
sites. Statistical methods then fit observed diversity to the expected level
under a particular theoretic model of linked selection processes. Consequently,
statistical inferences about selection may be incorrect when the underlying
theory inaccurately predicts average pairwise coalescence times.

@@

Given that Santiago and Caballero's (\citeyear{Santiago2016-mu}) model is a
haploid-model for a single region under equilibrium negative selection, we
first sought to confirm it could accurately predict the reduction in realistic
forward-time simulations of diploid populations with recombination. Using SLiM
\parencite{Haller2019-vu,Haller2023-uk} we simulated a region of $10^5$
basepairs under varying levels of mutation and selection (see Methods
\ref{sec:methods-segsim}). We find a close correspondence between the observed
and predicted reductions in effective population size $B=\nicefrac{N_d}{N}$
over all selection and mutation parameters including weak selection (Figure
\ref{fig:figure-1}A), in contrast to classic BGS theory. Furthermore, to
investigate whether this accuracy was caused by the model correctly predicting
the equilibrium fitness variance and ratchet rate, we also measured these
throughout the simulation. Again, we find the diploid SC16 theory accurately
predicts both the ratchet rate (Figure 1\ref{fig:figure-1}B) and the genic
fitness variance (Figure 1\ref{fig:figure-1}C).

@@

Our method takes tracks of annotated features (an ``annotation model") that are
\emph{a priori} expected to have similar fitness effects, and estimates the
distribution of fitness effects for each feature type. We consider two classes
of annotation models: (1) CADD score models, which assign a rank-order
pathogenicity score to each basepair, and (2) and more interpretable gene
feature-based models that includes protein coding regions, intron and UTRs, and
PhastCons regions. We include PhastCons regions because there is abundant
evidence of highly-conserved non-coding regions in humans under negative
selection
\parencite{Meader2010-hm,Harmston2013-tt,Katzman2007-gq,Siepel2005-wh}, but
would not be included in gene-based annotation. Since our method fits a DFE for
each feature type, PhastCons regions that overlap other gene annotations (e.g.
coding sequences) must be assigned to either of the two feature types. Since
this can impact estimates, we fit both alternatives: a PhastCons Priority
model, where genic features that overlap PhastCons regions are classified as
PhastCons, and a Feature Priority model, where gene structure features take
priority and the PhastCons type catches all highly-conserved non-genic regions.
In total, we fit four annotation models (CADD 6\%, CADD 8\%, PhastCons
Priority, and Feature Priority) to high-coverage 1000 Genome data for three
populations: Yoruba (YRI), Han Chinese (CHB), and European (CEU) ancestries. We
assess and compare our models according to how well they predict patterns of
diversity on whole chromosomes left-out during the model fitting process. We
use the metric $R_\text{LOO}^2$, which is the proportion of the observed
variance in genomic diversity at the megabase scale predicted by our model. We
experimented with a few smaller spatial scales, but our results were consistent
with previous results suggesting the human linked selection signal fits best at
the megabase scale \parencite{Murphy2022-sj}.

@@

Next, we developed a composite-likelihood method to estimate pairwise diversity
in the absence of linked selection $\pi_0$, the deleterious mutation rate
($\mu_D$), and distribution of fitness effects ($\mathbf{W}$) under negative
selection. Because the focus of our work is to fix deficiencies in the classic
BGS model, our current implementation does not consider reductions in diversity
due to recurrent hitchhiking or other positive selection processes. Since
previous work has established that purifying selection is the dominant mode of
linked selection in humans and hitchhiking plays a limited role
\parencite{McVicker2009-ax,Murphy2022-sj}, humans are a suitable organism to
test our method.

@@

@@

We use a binomial likelihood with probability $\bar{\pi}(b; \Psi) = \pi_0
\bar{B}(b; \mu, \mathbf{W})$, where bars indicate averages over some bin width.
With multiple features classes and a map $k(g)$ between a segment $g$ to its
class, the reduction map is,

@@

Here, $\mu$ is the mutation rate and $\mathbf{W}$ is the $n_s \times K$ DFE
matrix over the selection grid and features. Since each column of the DFE
$\mathbf{W}$ must sum to one, numeric optimization was done in unconstrained
$n_s-1$ dimensional softmax space. Our binomial likelihood modeled the number
of different pairwise comparisons observed per window given the predicted
pairwise diversity and total number of comparisons implied per window (see
Supplementary Materials Section \ref{supp:data-summary}).

We tested multiple different optimization routines, finding that BOBYQA
outperformed alternatives \parencite{Powell_undated-ec}. We inspected and
confirmed convergence with diagnostic plots finding stable optima across 10,000
random starts (see Supplementary Materials Section \ref{suppsec:optima}).

\subsection*{Segment Simulations}
\label{sec:methods-segsim}

To assess the accuracy of the SC16 theory, we conducted forward simulations
across grids of mutation rates ($\log_{10}(\mu) = \{10^{-9.5}, 10^{-9}, \ldots,
10^{-8}\}$) and selection coefficients ($\log_{10}(\mu) = \{-1, 1.5, 2, \ldots,
4.5\}$) using SLiM \parencite{Haller2023-uk,Haller2019-vu}. Each simulation
evolved a 100kb region with deleterious mutations entering at rate $\mu$
per-basepair, per generation, each with an additive selection coefficient of
$-s$. Total fitness was multiplicative across sites. Each simulation recorded
the number of fixations and the fitness variance per generation. After $10N$
generations, the ARG of the tree was recorded and downstream Python scripts
calculated diversity in the center of the region. We averaged the results of
$10,000$ replicate simulations to compare against theory.

\subsection*{Chromosome-wide Simulations}

, using the HapMap recombination map (XXX) and in coding regions, PhastCons
regions, and UTRs as an approximate set of putatively conserved segments

We calculate the mean squared error (MSE) and relative error between
whole-chromosome simulations and the predicted B and B' maps by averaging
branch diversity in 10 kilobase windows (tskit, XXX) and dividing by the
neutral expectation of $4N$. The theoretic lower bound of the MSE is due the
noise in estimating the expected reduction map from simulations. For $n=100$
replicates, the sampling noise is minimal relative to the evolutionary noise in
the coalescence process \parencite{Tajima1983-gu}. We use this as the theoretic
lower bound assuming the maximum coalescence noise at $B=1$, $\var(\bar{B})
\approx \nicefrac{2}{9n}$.

\subsection*{Human Population Genomic Diversity Estimates}

We estimate pairwise diversity for Yoruba (YRI), European (CEU), and Han
Chinese (CHB) ancestry individuals in the high-coverage 1000 Human Genomes
dataset (XXX). We

\subsection*{Local-rescaled }
We note that it is computationally intractable to solve this equation excluding
the contribution of the focal segment

\subsection*{Assessing the B' Maps}

We first assessed the B' maps against the observed average reduction across
$XXX$ forward-simulations, $\bar{B}_\text{sim}(x)$. The mean-squared error has
two sources, (1) bias from a breakdown in our theory $B(x)$ for the average
reduction, and (2) variance from the simulations. We show that for most of the
parameter space, the measured mean-squared error between simulation averages
and our B' maps is at the theoretic minimum. This minimum variance is entirely
from the coalescent noise when estimating $\bar{B}\text{sim}(x)$, which

\subsection*{Maximum Likelihood Estimation}

\subsection*{Local Rescaling and Substitution Rate Prediction}

>>>>

(c.f. \cite{Robertson1961-ho,Santiago1995-hx}; see Supplementary Section
\ref{supp:theory} for a proof). In general, however, statistics such as
heterozygosity depend on the cumulative autocorrelation factor up to some time
$t$, $Q_t^2$ \parencite[p. 2111]{Santiago1998-bs}. This reflects the fact that
effective population size experienced by a mutation newly associated with a
fitness background is larger than the asymptotic effective population size,
since not much autocorrelation has accumulated in the stochastic perturbations
caused by drift yet. From a backwards-time perspective, this reflects the fact
that pairwise diversity (which is approximately heterozygosity when $2N\mu \ll
1$) depends on the pairwise coalescence rate per generation, which is not
constant under draft. The benefit of using Robertson's forward-time model of
draft is that the inflation factor is invariant with respect to the particular
fitness background the neutral allele becomes stochastically associated with.
By contrast, the difficulty with modeling draft backwards in time is that the
coalescence rates experienced by a lineage are not invariant to which lineage
was sampled due to its particular associated fitness background.

@@

where the factor of one-half comes from ignoring weak associations from
unlinked regions and chromosomes (see Supplementary Section
\ref{supp:heritable-fitness}). In our genome-wide model, we consider the
summation in Equation \eqref{eq:polygenic_Ne} over non-overlapping segments $i
\in \{1, 2, \ldots, S\}$ each undergoing selection such that segment $i$
contributes additive fitness variance $V_{A,i}$ to the total additive genetic
fitness variance. Under equilibrium levels of fitness variation, the
autocorrelation function for a neutral allele associated with segment $i$ is
$C(t) = [(1-r_i)(1-\kappa_i)]^t$, where $r_i$ is the recombination fraction to
the segment and $\kappa_i$ is the rate that the associated fitness variance
decays due to selective dynamics. Then, the cumulative autocorrelation is,

@@

Similar to previous work, our method characterizes the negative selection
process through the signal left on nucleotide diversity at linked sites.
Although nucleotide diversity is a less-informative statistic than the full
site frequency spectrum, it is much easier to derive models of how selection
alters expected pairwise coalescence times at linked sites than to describe the
entire genealogy. Spatial patterns of pairwise diversity are highly variable
across eukaryotic genomes, as linked selection leads diversity to become
correlated with spatial heterogeneity in recombination rates and the density of
conserved sites. Statistical methods then fit observed diversity to the
expected level under a particular theoretic model of linked selection
processes. Consequently, statistical inferences about selection may be
incorrect when the underlying theory inaccurately predicts average pairwise
coalescence times.

@@

Given that Santiago and Caballero's (\citeyear{Santiago2016-mu}) model is a
haploid-model for a single region under equilibrium negative selection, we
first sought to confirm it could accurately predict the reduction in realistic
forward-time simulations of diploid populations with recombination. Using SLiM
\parencite{Haller2019-vu,Haller2023-uk} we simulated a region of $10^5$
basepairs under varying levels of mutation and selection (see Methods
\ref{sec:methods-sim}). We find a close correspondence between the observed
and predicted reductions in effective population size $B=\nicefrac{N_d}{N}$
over all selection and mutation parameters including weak selection (Figure
\ref{fig:figure-1}A), in contrast to classic BGS theory. Furthermore, to
investigate whether this accuracy was caused by the model correctly predicting
the equilibrium fitness variance and ratchet rate, we also measured these
throughout the simulation. Again, we find the diploid SC16 theory accurately
predicts both the ratchet rate (Figure 1\ref{fig:figure-1}B) and the genic
fitness variance (Figure 1\ref{fig:figure-1}C).

@@

Our method takes tracks of annotated features (an ``annotation model") that are
\emph{a priori} expected to have similar fitness effects, and estimates the
distribution of fitness effects for each feature type. We consider two classes
of annotation models: (1) CADD score models, which assign a rank-order
pathogenicity score to each basepair, and (2) and more interpretable gene
feature-based models that includes protein coding regions, intron and UTRs, and
PhastCons regions. We include PhastCons regions because there is abundant
evidence of highly-conserved non-coding regions in humans under negative
selection
\parencite{Meader2010-hm,Harmston2013-tt,Katzman2007-gq,Siepel2005-wh}, but
would not be included in gene-based annotation. Since our method fits a DFE for
each feature type, PhastCons regions that overlap other gene annotations (e.g.
coding sequences) must be assigned to either of the two feature types. Since
this can impact estimates, we fit both alternatives: a PhastCons Priority
model, where genic features that overlap PhastCons regions are classified as
PhastCons, and a Feature Priority model, where gene structure features take
priority and the PhastCons type catches all highly-conserved non-genic regions.
In total, we fit four annotation models (CADD 6\%, CADD 8\%, PhastCons
Priority, and Feature Priority) to high-coverage 1000 Genome data for three
populations: Yoruba (YRI), Han Chinese (CHB), and European (CEU) reference
samples. We assess and compare our models according to how well they predict
patterns of diversity on whole chromosomes left-out during the model fitting
process. We use the metric $R_\text{LOO}^2$, which is the proportion of the
observed variance in genomic diversity at the megabase scale predicted by our
model. We experimented with a few smaller spatial scales, but our results were
consistent with previous results suggesting the human linked selection signal
fits best at the megabase scale \parencite{Murphy2022-sj}.

@@

Next, we developed a composite-likelihood method to estimate expected diversity
without linked selection $\pi_0$, the deleterious mutation rate ($\mu_D$), and
distribution of fitness effects ($\mathbf{W}$) under negative selection.
Because the focus of our work is to fix deficiencies in the classic BGS model,
our current implementation does not consider reductions in diversity due to
recurrent hitchhiking or other positive selection processes. Since previous
work has established that purifying selection is the dominant mode of linked
selection in humans and hitchhiking plays a limited role
\parencite{McVicker2009-ax,Murphy2022-sj}, humans are a suitable organism to
test our method.

@@

Regularization / overfitting

@@

We use a binomial likelihood with probability $\bar{\pi}(b; \Psi) = \pi_0
\bar{B}(b; \mu, \mathbf{W})$, where bars indicate averages over some bin width.
With multiple features classes and a map $k(g)$ between segments $g$ to their
class, the reduction map is,

@@

Here, $\mu$ is the mutation rate and $W_{j,k(g)}$ is a DFE entry in the
$\mathbf{W}$ is the $n_s \times K$ DFE matrix over the selection grid and
features. Since each column of the DFE $\mathbf{W}$ must sum to one, numeric
optimization was done in unconstrained $n_s-1$ dimensional softmax space. Our
binomial likelihood modeled the number of different pairwise comparisons
observed per window given the predicted pairwise diversity and total number of
comparisons implied per window (see Supplementary Materials Sections
\ref{supp:data-summary} and \ref{supp:likelihood}).

We tested multiple different optimization routines, finding that BOBYQA
outperformed alternatives \parencite{Powell_undated-ec}. We inspected and
confirmed convergence with diagnostic plots finding stable optima across 10,000
random starts (see Supplementary Materials Section \ref{suppsec:optima}). We
assessed model fit using out-sample predictive error, calculated by leaving out
a whole chromosome during fitting and predicting its diversity. To calculate
uncertainty, we used a block jackknife approach in 10 Mbp windows
(Supplementary Section \ref{supp:jackknife}).

\subsection*{Human Population Genomic Data}
\label{sec:methods-data}

Our analyses was conducted on the Yoruba (YRI), European (CEU), and Han Chinese
(CHB) reference samples individuals from the high-coverage 1000 Human Genomes
data aligned to GRCh38/hg38 \parencite{Byrska-Bishop2022-tn}. Since nucleotide
diversity is a ratio estimator, it can be biased when subtly different
filtering criteria are applied to variant and invariant sites. To prevent this,
we conducted our analyses on Genomic VCF (gVCF) files that contain genotype
calls for both variant and invariant sites \parencite{Illumina_Inc2020-dk}.
Then, we apply the same genotype filtering criteria to all called sites
(Supplementary Materials Section \ref{supp:filter}). We also applied sequence
accessibility masks that containing only non-repeat, non-centromeric sequence
that passed the 1000 Genomes strict filter
(\cite{1000_Genomes_Project_Consortium2015-mi}; Supplementary Material Section
\ref{supp:accessible}). Since our theory only considers the indirect effects of
linked selection on a site, we additionally masked sites that are likely under
direct selective constraint (see Supplementary Materials Section
\ref{supp:neutral}). Finally, for every basepair passing these filtering and
masking criteria, we counted the number of reference and alternative allele
counts (excluding all multiallelic, indel, and CNV variants).

For all of our main analyses, we used the recombination map from
\textcite{Halldorsson2019-ey} estimated from a trio-based design to avoid
circularity that could occur by using LD-based maps. We use Ensembl gene
annotation \parencite{Cunningham2022-vk}, a special CADD Score dataset with
McVicker B scores removed (to avoid circularity;
\cite{Kircher2014-bv,Rentzsch2019-lr}), and PhastCons regions
\parencite{Siepel2005-wh}.

\subsection*{Forward Simulations}
\label{sec:methods-sim}

We use two types of forward simulations in this paper; both use SLiM version
4.0.1 \parencite{Haller2023-uk,Haller2019-vu}. First, we simulated a 100 kbp
segment in a population of $N=1000$ diploids across grids of mutation and
selection parameters to assess the accuracy of the SC16 theory. Each simulation
recorded the number of fixations and the fitness variance per generation, and
at the end $10N$ generations, the ancestral recombination graph was recorded so
downstream Python scripts could calculate diversity. Each of these segment
simulations was replicated $10,000$ times and averaged to compare against
theoretic expectations.

Second, we conducted forward simulations of negative selection on whole human
chromosomes. These simulations used the HapMap recombination map
\parencite{International_HapMap_Consortium2007-nu} and treated coding,
PhastCons, and UTR regions as an approximate set of conserved regions. We used
these whole-chromosome simulations for three purposes. First, we simulated
negative selection on chromosome 10 to confirm our method to estimate classic B
and new B' maps matched the average empirical reduction map across a grid of
mutation and selection parameters. Second, we evaluated our composite
likelihood method by simulating negative selection on the first five human
chromosomes, across grids of mutation and selection parameters. Multiple
independently-evolved chromosomes were combined into a synthetic genome in
order to have enough data for our procedure. Third, we conducted additional
simulations like these, but varying the dominance coefficient and expanding the
population by a factor of 1.004 each generation, starting $9.3$ generations in
to mimic the human expansion out-of-Africa (XXX). More details about these
simulations can be found in Supplementary Section \ref{supp:simulation}.

### 2023-07-01 manuscript/supplementary.tex
more methods

\section{Quantitative Genetic Linked Selection Model Theory}
\label{suppsec:theory}

@@

\subsection{Heritable Fitness and the Accumulation of Autocovariance}
\label{suppsec:heritable-fitness}

@@

\subsection{Pairwise Diversity Estimates}
\label{suppsec:diversity}

@@

\subsection{}

In order to try to ensure accurate estimation of pairwise diversity, which is a
ratio estimator that is sensitivity to its denominator, we use the complete
per-basepair genotype calls (gVCF) produced by Illumina's DRAGEN pipeline
\parencite{Illumina_Inc2020-dk}. The original samples were from 178 Yoruban
individuals sequenced to 30x by the New York Genome Center
\parencite{Byrska-Bishop2022-tn}. This allows for filtering to be applied to
the entire genome at once, rather than just variants, so the denominator does
not need to be estimated separately.

@@

\subsection{Filtering gVCFs}

@@

  \item \textbf{Reference masking}: soft and hard-masked regions in the human
      GRCh38 reference genome are also masked. Soft-masked regions were
      determined by Ensembl (XXX), which uses Repeat Masker (XXX) and Dust
      (XXX). Additionally, we used the 1000 Genomes strict filter (XXX).

@@

\subsection{Putatively Neutral Tracks}

@@

where $V_\text{res}$ is the mean squared error across windows and
$V_\text{tot}$ is the total variance in the predicator. Suppose we have
estimates $\widehat{\pi}_0 \widehat{B}_i$ for $\pi_0 B_i$ from our model. Then,

@@

The total variance in the predictor is,

@@

\subsection{Likelihood}

@@

As in Section \ref{suppsec:diversity}, the number of pairwise differences and
the total number of pairwise comparison at a site are sufficient statistics for
the likelihood. Then, the binomial log-likelihood for the data at site $v$ is,

@@

\subsection{Out-Sample Error Estimation and $R_\text{LOO}^2$}

We fit each of our B' model fits 22 times, excluding a whole chromosome each
time to calculate out-sample prediction error. Each fit used 10,000 random
starting positions as the main fits. The out-sample prediction error for all
bins in chromosome $C$ left out is,

\begin{align}
    SSE(C) = \sum_{b \in C} (\widehat{\pi}_{b,Y\not \in C}  - \pi_b)^2
\end{align}
%

where $\widehat{\pi}_{b,Y\not \in C}$ denotes the predictions for window $b$
fit on data excluding chromosome $C$. The average out-sample error $SSE$ across
chromosomes is then the average of $SSE(C)$ across chromosomes, weighted by the
number of bins per chromosome. We then calculate the out-sample coefficient of
determination $R_\text{LOO}^2$

\begin{align}
    R_\text{LOO}^2 = \frac{SSE}{}
\end{align}

\subsection{Jackknife Procedures}

@@

@@

\subsection{Optimization and Diagnostics}
\label{suppsec:optim}

>>>>

\section{Quantitative Genetic Linked Selection Model Theory}
\label{supp:theory}

@@

\subsection{Heritable Fitness and the Accumulation of Autocovariance}
\label{supp:heritable-fitness}

@@

\subsection{Pairwise Diversity Estimates}
\label{supp:diversity}

@@

In order to try to ensure accurate estimation of pairwise diversity, which is a
ratio estimator that is sensitivity to its denominator, we use the complete
per-basepair genotype calls (gVCF) produced by Illumina's DRAGEN pipeline
\parencite{Illumina_Inc2020-dk}. The original samples were sequenced to 30x by
the New York Genome Center \parencite{Byrska-Bishop2022-tn}. This allows for
filtering to be applied to the entire genome at once, rather than just
variants, so the denominator does not need to be estimated separately.

@@

\subsection{Filtering gVCFs}
\label{supp:filter}

@@

  \item \textbf{Reference masking}: soft and hard-masked regions in the human
      GRCh38 reference genome are also masked. Soft-masked regions were
      determined by Ensembl \parencite{Cunningham2022-vk}, which uses Repeat
      Masker \parencite{Smit2015-nm} and Dust \parencite{Morgulis2006-uo}. We
      also only used sites that pass the 1000 Genomes strict filter
      \parencite{1000_Genomes_Project_Consortium2015-mi}.

@@

\subsection{Putatively Neutral Tracks}
\label{supp:neutral}

@@

where $V_\text{res}$ is the mean squared error across windows and
$V_\text{tot}$ is the total variance in the observed values. Suppose we have
estimates $\widehat{\pi}_0 \widehat{B}_i$ for $\pi_0 B_i$ from our model. Then,

@@

The total variance among the observed diversity values is,

@@

\subsection{Likelihood}
\label{supp:likelihood}

@@

As in Section \ref{supp:diversity}, the number of pairwise differences and
the total number of pairwise comparison at a site are sufficient statistics for
the likelihood. Then, the binomial log-likelihood for the data at site $v$ is,

@@

\subsection{Out-Sample Error Estimation and $\widehat{R}_\text{LOCO}^2$}

To estimate out-sample prediction error, we fit each model leaving one
chromosome out (LOCO) and calculating the out-sample prediction error on
these excluded chromosomes. Each fit used 10,000 random starting positions as
the main fits. The out-sample prediction error (i.e. sum of residuals
squared, $\widehat{\mathrm{SSR}}_\text{LOCO}$) for all bins in chromosome $C$
left out is,

\begin{align}
    \widehat{\mathrm{SSR}}_\text{LOCO}(C) = \sum_{b \in C} (\widehat{\pi}_{b, (C)}  - \pi_b)^2
\end{align}
%
where $\widehat{\pi}_{b,(C)}$ denotes the predictions for window $b$ fit on
data excluding chromosome $C$. The average out-sample error
$\widehat{\mathrm{SSR}}_\text{LOCO}$ across chromosomes is then the average of
$\widehat{\mathrm{SSR}}_\text{LOCO}(C)$ across chromosomes, weighted by the
number of bins per chromosome. We then calculate the out-sample coefficient of
determination $\widehat{R}_\text{LOCO}^2$

\begin{align}
    \widehat{R}_\text{LOCO}^2 = 1-\frac{\widehat{\mathrm{SSR}}_\text{LOCO}}{\widehat{\mathrm{SST}}_\text{LOCO}}
\end{align}
%
where,

\begin{align}
    \widehat{\mathrm{SST}}_\text{LOCO}(C) = \sum_{b \in C} (\widehat{\pi}_{b, (C)}  - \bar{\pi})^2
\end{align}
%
and $\bar{\pi}$ is the average diversity across all windows.

\subsection{Jackknife Procedures}
\label{supp:jackknife}

@@

\subsection{Chromosome 10 Simulations}

XXX
, using the HapMap recombination map (XXX) and in coding regions, PhastCons
regions, and UTRs as an approximate set of putatively conserved segments

We calculate the mean squared error (MSE) and relative error between
whole-chromosome simulations and the predicted B and B' maps by averaging
branch diversity in 10 kilobase windows (tskit, XXX) and dividing by the
neutral expectation of $4N$. The theoretic lower bound of the MSE is due the
noise in estimating the expected reduction map from simulations. For $n=100$
replicates, the sampling noise is minimal relative to the evolutionary noise in
the coalescence process \parencite{Tajima1983-gu}. We use this as the theoretic
lower bound assuming the maximum coalescence noise at $B=1$, $\var(\bar{B})
\approx \nicefrac{2}{9n}$.

@@

\subsection{Optimization and Diagnostics}
\label{supp:optim}
